Variation on a Variational Principle

### 2020-10-05 draft1.tex
done! hopefully...

\documentclass[11pt]{amsart}
\usepackage{float}
\usepackage{amsfonts, amstext, amsmath, amsthm, amscd, amssymb, upgreek}
\usepackage{bbm}
\usepackage{graphicx, color,  subfigure, wrapfig, overpic}
%\usepackage[all,cmtip]{xy} 
\usepackage{tikz}
%\usepackage{mystyle}

\newcommand{\thmref}[1]{Theorem \ref{#1}}
\newcommand{\prpref}[1]{Proposition \ref{#1}}
\newcommand{\lemref}[1]{Lemma \ref{#1}}
%\newcommand{\ref}[1]{Figure \ref{#1}}
\newcommand{\comment}[1]{}

@@

\section{Introduction}

@@

Menasco \cite{Menasco} showed there are decompositions of the complements of 
alternating links in $\Sp^3$ into two polyhedra, a top polyhedron and a bottom polyhedron. 
For alternating links, $K$ in the thickened torus Champanerkar, Kofman and Purcell \cite{CKP2}
showed that there is a decomposition of the complement of $K$ into objects called torihedra, which we think of 
as counterparts of Menasco's decomposition of links in $\Sp^3$ into polyhedra, for links in the thickened 
torus. Just like Menasco's decomposition, one can obtain a top and bottom torihedron. In section 2 we show that for augmented links in the thickened torus 
one can also obtain a decomposition of the complement into a top and bottom torihedron.

@@

 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\section{Augmented Links}
 Champanerkar, Kofman and Purcell have studied alternating links in the thickened torus \cite{CKP2}. They define a link in the thickened torus as a quotient of a biperiodic alternating link as follows,

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The following figures is an example which decomposes the link (C) of Figure
\ref{fig:Augmentations}. 
 \begin{figure}[h] 
 \centering
 \includegraphics
[height=3cm]{fig-3}
 \includegraphics [height=3cm]{fig-4}
  \caption{Each crossing
circle bounds a disk} \end{figure}

 \vspace*{-0.5cm} 
 \begin{figure}[h] 
 \centering 
 \includegraphics
 [height=3cm]{fig-5} 
 \caption{We split the disk and collapse the arc of each
 crossing circle to ideal vertices} 
 \end{figure}

\vspace*{-0.5cm} 
\begin{figure}[h] 
\centering 
\includegraphics[height=3cm]{fig-6} 
\caption{Left: The crossing arc is the edge in red.
Middle: Picture of splitting the crossing edge. Right: The link component is
pushed off to infinity.} 
\end{figure}

 \vspace*{-0.5cm} 
 \begin{figure}[h] 
 \centering 
 \includegraphics
 [height=3cm]{top-bottom} 
 \caption{Left: The top torihedron. Right: The bottom torihdron with rotation for face gluing.} 
 \label{fig:top-bottom} \end{figure}

\comment{
\begin{remark}\label{cor:triangulation}
The faces of one torihedron which do not correspond to a bow-tie glue to a
corresponding face of the second torihedron to form bipyramids. One can add
stellating edges, edges which have vertices corresponding to the different
components of the Hopf link that cuts through the center of the face to
decompose the bipyramids into tetrahedra. This process is called {\it
stellation}. Stellating all the bipyramids of the decomposition obtained from
Proposition \ref{p:tori_decomp} we obtained a triangulation of the
complement of $L$ which is made up of {\it horizontal} edges (edges from the
graph of the torihedra), {\it vertical} edges (edges from coning vertices of the
the graph of the torihedra) and stellating edges.
\end{remark}
}

\comment{
\begin{proof}[Proof of Remark 2.7]
Using the decomposition from Proposition
\ref{p:tori_decomp} we can add a stellation edge whose end points
are $\torus \times \{-1\}$ and $\torus \times \{1\}$ which cut through faces of $\torus \times I$ which do not come from the augmentation disks like in \cite{CKP2}.
Since each face of the top torihedron gets glued to the bottom torihedron, we
obtain bipyramids, and the stellation edge will decompose the bipyramids into
tetrahedra. Then the link of the vertex  $\torus \times \{1\}$ or
$\torus \times \{-1\}$ is the graph of the torihedron
%$\torus \times [0,1)$ or
%$\torus \times (-1,0]$ respectively.
\end{proof}
}

TODO refer to BS about theta's
\begin{definition}
An \emph{angled torihedron} $(\sT, \theta_\bullet^*)$
is a torihedron $\sT$ with
an assignment $\theta_e^* \in [0,\pi]$
such that for each vertex $v \in G(\sT)$,
$\sum_{e \ni v} \theta_e^* = (\deg(v) - 2)\pi$.
We also denote $\theta_e = \pi - \theta_e^*$,
so $\sum_{e \ni v} \theta_e = 2\pi$;
we refer to $\theta_e$ as the exterior angle
and $\theta_e$ as the interior angle.

@@

TODO same as [BS] ``coherent angle system'' or something.

TODO remark on what splitting means in terms of ideal hyperbolic

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Hyperbolicity of Augmented Links}
Thurston introduced a method for finding the unique hyperbolic metric for a given 3-manifold $M$ with boundary consisting of tori \cite{Thurston}. The idea was to triangulate the interior of $M$ into ideal tetrahedra and give those tetrahedra hyperbolic shapes (called shape parameters) that glue up coherently in $M$. The shape parameter of a tetrahedron is described by the cross-ratio of its four vertices on the sphere at infinity. Thurston had written down a system of gluing equations with shape parameters whose solutions correspond to the complete hyperbolic metric on the interior of $M$. Casson and Rivin separated gluing equations into a linear and non-linear part \cite{Casson-Rivin}. Angle structures is the linear part of Thurston's gluing equations, and what we will use to attain hyperbolicity of complements of augmented links in the thickened torus.

@@

For a hyperbolic link $K$ in $\torus \times I$, we show
that the resulting link obtained from augmenting $K$ is
hyperbolic. The idea is to start with a graph from the torihedral decomposition
of the link $K$ which will give us a graph on each torihedron with $\pi/2$ edges
\cite{CKP2},
the angled torihedra can be further decomposed into base-angled pyramids.
By \prpref{p:tori_decomp},
there is a torihedral decomposition of the complement of the augmented link $L$.
Using those angles,
we then assign angles to edges of these torihedra
and decompose them into base-angled pyramids.
These pyramids can be further decomposed into tetrahedra,
thus obtaining an angle structure on a triangulation.

@@

\begin{theorem}
Let $K$ be a weakly prime, alternating link whose diagram has no bigons.
Let $L$ be a link obtained from augmenting $K$.
Then $L$ is hyperbolic.
\end{theorem}

@@

%Thus, we could attempt to apply the feasible flow theorem in a similar manner,
%and obtain an angle-splitting for $\sT_T(L)'$, but 
TODO rephrase nicely.
We can directly get an angle-splitting for $\sT_T(L)'$
using the angle-splitting for $\sT_T(L)$.
We reuse the $+/-$ assignments from Figure \ref{f:adding_edges}.
For $e = e_{i,0}$, we increase $\vphi_{\vec{e}}, \vphi_{\cev{e}}$
by $\veps'/2$ each; for $e = e_{i,\pi}$,
we decrease them by $\veps'/2$ each.
For other edges, we increase/decrease $\vphi_{\vec{e}}$ by $\veps'$,
where $\vec{e}$ is the oriented edge corresponding to the side
on which the $+/-$ sign appears in Figure \ref{f:adding_edges};
so for example, if an edge $e$ receives both $+$ and $-$,
then one of $\vphi_{\vec{e}},\vphi_{\cev{e}}$ increases
while the other decreases, thus $\theta_e^*$ remains constant.

@@

We need to check that upon gluing $\sT_T(L)'$ to $\sT_B(L)'$,
the sum of dihedral angles around each edge is still $2\pi$.
This was clearly true before deforming, as the new edges of
$\Gamma_T(L)'$ only gets identified with the unique
corresponding edge of $\Gamma_B(L)'$, and they are both labeled
with $\theta^* = \pi$.
To see that the deformation does not change these sums,
note that in the identification of faces of $\Gamma_T(L)'$
to $\Gamma_B(L)'$,
an edge with $\theta^*=0$ is identified with an edge with
$\theta^*=\pi$,
so an increase in the former would be conterbalanced by
a decrease in the latter.
It is also easy to see this is the case for the other edges.
TODO hmm more detail?

>>>>

\documentclass[11pt]{amsart}
\usepackage{float}
\usepackage{amsfonts, amstext, amsmath, amsthm, amscd, amssymb, upgreek}
\usepackage{bbm}
\usepackage{graphicx, color,  subfigure, wrapfig, overpic}
\usepackage{fullpage}
%\usepackage[all,cmtip]{xy} 
%\usepackage{tikz}
\usepackage{tikz-cd}
%\usepackage{mystyle}

\newcommand{\thmref}[1]{Theorem \ref{#1}}
\newcommand{\prpref}[1]{Proposition \ref{#1}}
\newcommand{\lemref}[1]{Lemma \ref{#1}}
\newcommand{\figref}[1]{Figure \ref{#1}}
\newcommand{\secref}[1]{Section \ref{#1}}
\newcommand{\eqnref}[1]{(\ref{#1})}
%\newcommand{\ref}[1]{Figure \ref{#1}}
\newcommand{\comment}[1]{}
\newcommand{\inv}{{-1}} % annoying to type {-1} when taking inverse

@@

\section{Introduction}
\label{s:intro}

@@

Menasco \cite{Menasco} showed there are decompositions of the complements of 
alternating links in $\Sp^3$ into two polyhedra, a top polyhedron and a bottom polyhedron. 
For alternating links, $K$ in the thickened torus Champanerkar, Kofman and Purcell \cite{CKP2}
showed that there is a decomposition of the complement of $K$ into objects called torihedra,
which we think of 
as counterparts of Menasco's decomposition of links in $\Sp^3$ into polyhedra, for links in the thickened 
torus. Just like Menasco's decomposition, one can obtain a top and bottom torihedron.
In \secref{s:auglinks} we show that for augmented links in the thickened torus 
one can also obtain a decomposition of the complement into a top and bottom torihedron.
In \secref{s:hyperbolicity}, we prove that augmented alternating links in the thickened torus
are hyperbolic.
In the Appendix, we provide an alternative proof to hyperbolicity of these links
via circle patterns, which one could use obtain lower bounds for volume.

@@

 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\section{Augmented Links}
\label{s:auglinks}

Champanerkar, Kofman and Purcell have studied alternating links in the thickened torus \cite{CKP2}. They define a link in the thickened torus as a quotient of a biperiodic alternating link as follows,

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The following figures is an example which decomposes the link (C) of Figure
\ref{fig:Augmentations}.

\begin{figure}[h] 
\centering
\includegraphics[height=3cm]{fig-3}
\includegraphics[height=3cm]{fig-4}
	\caption{Each crossing circle bounds a disk}
\end{figure}

%\vspace{-0.5cm} 
\begin{figure}[h] 
\centering 
\includegraphics[height=3cm]{fig-5} 
	\caption{We split the disk and collapse the arc of each
 crossing circle to ideal vertices} 
\end{figure}

\begin{figure}[h] 
\centering 
\includegraphics[height=3cm]{fig-6} 
\caption{Left: The crossing arc is the edge in red.
Middle: Picture of splitting the crossing edge. Right: The link component is
pushed off to infinity.} 
\end{figure}

\begin{figure}[h] 
\centering 
\includegraphics[height=3cm]{top-bottom} 
	\caption{Left: The top torihedron. Right: The bottom torihdron with rotation for face gluing.} 
\label{fig:top-bottom}
\end{figure}

\begin{definition}
An \emph{angled torihedron} $(\sT, \theta_\bullet^*)$
is a torihedron $\sT$ with
an assignment $\theta_e^* \in [0,\pi]$
such that for each vertex $v \in G(\sT)$,
$\sum_{e \ni v} \theta_e^* = (\deg(v) - 2)\pi$.
We also denote $\theta_e = \pi - \theta_e^*$,
so $\sum_{e \ni v} \theta_e = 2\pi$;
we refer to $\theta_e$ as the exterior angle
and $\theta_e$ as the interior angle.

@@

\begin{remark}
These $\theta$'s are related to the $\theta$'s in
\cite{BandS},
and the $\vphi$'s are related to their ``coherent angle system''.
\end{remark}

%TODO remark on what splitting means in terms of ideal hyperbolic

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Hyperbolicity of Augmented Links}
\label{s:hyperbolicity}
Thurston introduced a method for finding the unique hyperbolic metric for a
given 3-manifold $M$ with boundary consisting of tori \cite{Thurston}. The idea
was to triangulate the interior of $M$ into ideal tetrahedra and give those
tetrahedra hyperbolic shapes (called shape parameters) that glue up coherently
in $M$. The shape parameter of a tetrahedron is described by the cross-ratio of
its four vertices on the sphere at infinity. Thurston had written down a system
of gluing equations with shape parameters whose solutions correspond to the
complete hyperbolic metric on the interior of $M$. Casson and Rivin separated
gluing equations into a linear and non-linear part \cite{Casson-Rivin}. Angle
structures is the linear part of Thurston's gluing equations, and what we will
use to attain hyperbolicity of complements of augmented links in the thickened
torus.

@@

For a hyperbolic link $K$ in $\torus \times I$, we show
that the resulting link obtained from augmenting $K$ is
hyperbolic. The idea is to start with a graph from the torihedral decomposition
of the link $K$ which will give us a graph on each torihedron with $\pi/2$ edges
\cite{CKP2},
the angled torihedra can be further decomposed into base-angled pyramids.
By \prpref{p:tori_decomp},
there is a torihedral decomposition of the complement of the augmented link $L$.
Using those angles,
we then assign angles to edges of these torihedra
and decompose them into base-angled pyramids.
These pyramids can be further decomposed into tetrahedra,
thus obtaining an angle structure on a triangulation.
In the Appendix, we show that this triangulation
also has an angle structure that
satisfies Thurston's gluing equations.

@@

\begin{theorem}
\label{t:auglink_hyp}
Let $K$ be a weakly prime, alternating link whose diagram has no bigons.
Let $L$ be a link obtained from augmenting $K$.
Then $L$ is hyperbolic.
\end{theorem}

@@

%Thus, we could attempt to apply the feasible flow theorem in a similar manner,
%and obtain an angle-splitting for $\sT_T(L)'$, but 
%TODO rephrase nicely.
We can directly get an angle-splitting for $\sT_T(L)'$
using the angle-splitting for $\sT_T(L)$.
We reuse the $+/-$ assignments from Figure \ref{f:adding_edges}.
For $e = e_{i,0}$, we increase $\vphi_{\vec{e}}, \vphi_{\cev{e}}$
by $\veps'/2$ each; for $e = e_{i,\pi}$,
we decrease them by $\veps'/2$ each.
For other edges, we increase/decrease $\vphi_{\vec{e}}$ by $\veps'$,
where $\vec{e}$ is the oriented edge corresponding to the side
on which the $+/-$ sign appears in Figure \ref{f:adding_edges};
so for example, if an edge $e$ receives both $+$ and $-$,
then one of $\vphi_{\vec{e}},\vphi_{\cev{e}}$ increases
while the other decreases, thus $\theta_e^*$ remains constant.

@@

We need to check that upon gluing $\sT_T(L)'$ to $\sT_B(L)'$,
the sum of dihedral angles around each edge is still $2\pi$.
This was clearly true before deforming, as the new edges of
$\Gamma_T(L)'$ only gets identified with the unique
corresponding edge of $\Gamma_B(L)'$, and they are both labeled
with $\theta^* = \pi$.
To see that the deformation does not change these sums,
note that in the identification of faces of $\Gamma_T(L)'$
to $\Gamma_B(L)'$,
an edge with $\theta^*=0$ is identified with an edge with
$\theta^*=\pi$,
so an increase in the former would be conterbalanced by
a decrease in the latter.
It is also easy to see this is the case for the other edges.
%TODO hmm more detail?

### 2020-10-05 notation.tex
done! hopefully...

A circle pattern is determined by the radius of the circle $C_f$ associated to each face, $r_f$,
and the angle that each edge subtends in adjacent faces, $\vphi_{\vec{e}}$.
(See \figref{f:circle_pattern})
\begin{figure}
\includegraphics[width=5cm]{circle_pattern}
\caption{Circle pattern in terms of radii $r_f$ and angles $\vphi_{\vec{e}}$}
\label{f:circle_pattern}
\end{figure}
Thus determines, and is determined by a point in $\RRR \times \QQQ$, where

@@

[TODO Check this] The usual notion of circle pattern would be restricted to those $c\in \CCC$
with $\vphi_{\vec{e}} \in (0,\pi)$ and $\theta_e \in (0,\pi)$.
One may consider deforming a circle pattern so as to have some $\theta_e$ approach 0 or $\pi$.
The limit $\theta_e \to \pi$ is easy to picture, one simply gets that the two circles
$C_f, C_{f'}$ of the adjacent faces become tangent. The limit $\theta_e \to 0$
is a bit more complex, as the final shape of $Q_e$ (the quadrangle associated to $e$)
may depend on $\vphi_{\vec{e}}$ or $\vphi_{\cev{e}}$.
If we parametrize circle patterns by $(r_f)$ and $(\theta_e)$ as in TODO \ocite{BS},
the limiting shape would depend on the relationship between 
$r_f$ and $r_{f'}$.
As our main result depends on extending \ocite{BS} result
to such limits and beyond,
we find it more convenient to describe extended circle patterns by
$(r_f)$ and $(\vphi_{\vec{e}})$.

@@

TODO non-degenerate circle pattern?

@@

%TODO figures for convex, thin faces

@@

\subsection{Modifications to Extended Circle Patterns}

Let $\Gamma$ be a graph on the torus, and let $c$ be a
non-singular extended circle pattern on $\Gamma$.
There are two simple operations to modify $\Gamma$:
cutting a face along a new edge,
and splitting a vertex in two then joining them by a new edge.
We describe how to associate an extended circle pattern $c'$ to
the new graph.

Cutting a face: Consider a face $f$,
and suppose we added a new edge $e'$ so that $f$ is split into
two faces $f'$ and $f''$, so that
$\del f' = \{\vec{e_1},\ldots,\vec{e_k}, \vec{e'}\}$,
and $\del f'' = \{\cev{e'}, \vec{e_{k+1}},\ldots \vec{e_n}\}$.
Then for the new extended circle pattern $c'$,
we set $r_{f'}(c') = r_{f''}(c') = r_f(c)$,
$\vphi_{\vec{e'}} = \pi - \vphi_{\vec{e_1}} - \ldots - \vphi_{\vec{e_k}}$,
$\vphi_{\cev{e'}} = \pi - \vphi_{\vec{e_{k+1}}} - \ldots \vphi_{\vec{e_n}}$,
and $r_\bullet(c') = r_\bullet(c)$, $\vphi_\bullet(c') = \vphi_\bullet(c)$
for all other faces, edges.
It is straightforward to check that $c'$ is non-singular.

Consider a vertex $v$ of the TODO

>>>>

A circle pattern is determined by the radius of the circle $C_f$ associated to each face, $r_f$,
and the angle that each edge subtends in adjacent faces, $\vphi_{\vec{e}}$.
(See \figref{f:circle_pattern})
\begin{figure}
\includegraphics[width=8cm]{circle_pattern}
\caption{Circle pattern in terms of radii $r_f$ and angles $\vphi_{\vec{e}}$}
\label{f:circle_pattern}
\end{figure}
Thus determines, and is determined by a point in $\RRR \times \QQQ$, where

@@

The notion of circle pattern that is considered in \cite{BandS}
would be restricted to those $c\in \CCC$
with $\vphi_{\vec{e}} \in (0,\pi)$ and $\theta_e \in (0,\pi)$.
They parametrize circle patterns by $(r_f)$ and $(\theta_e)$
(or $(\theta_e^*)$),
and call $(\vphi_{\vec{e}})$ a coherent angle system.
One may consider deforming a circle pattern so as to have
some $\theta_e$ approach 0 or $\pi$.
The limit $\theta_e \to \pi$ is easy to picture, one simply gets that the two circles
$C_f, C_{f'}$ of the adjacent faces become tangent. The limit $\theta_e \to 0$
is a bit more complex, as the final shape of $Q_e$ (the quadrangle associated to $e$)
may depend on $\vphi_{\vec{e}}$ or $\vphi_{\cev{e}}$.
If we parametrize circle patterns by $(r_f)$ and $(\theta_e)$ as in \cite{BandS},
the limiting shape would depend on the relationship between 
$r_f$ and $r_{f'}$.
As our main result depends on extending \ocite{BS} result
to such limits and beyond,
we find it more convenient to describe extended circle patterns by
$(r_f)$ and $(\vphi_{\vec{e}})$.

@@

@@

%figures for convex, thin faces

@@

### 2020-10-05 results.tex
done! hopefully...

TODO: make a list of images of vectors, e.g.
$\Theta_*(\ddd{\vphi_{\vec{e}}}) = -\ddd{\theta_e}$.

@@

To that end, let $L$ be the $(|E|-1)$-plane distribution on $\TTT$
tangent to the level sets of $\sum_{e\in E} \theta_e$.
The following proposition proves it (TODO avoid using `it'?) up to first order at a point:

\begin{proposition}
\label{p:point_lift}
Let $c \in \CCC$ be a
non-singular extended circle pattern on a graph
with no bigons or self-loops such that
there is no wide cycle of thin faces.
%\begin{itemize}
	%\item $\Phi_f = 2\pi$ for all $f\in F$;
	%\item there is no wide cycle of thin faces.
	%\item all faces are eithe convex or thin,
		%with at least one convex face;
	%\item every pair of faces is connected by a wide path.
%\end{itemize}
Let $K_c = \ker (\Phi_*|_c : T_c \CCC \to T_{\Phi(c)} \PPP)$.
Then $\Theta_*(K_c) = L_{\Theta(c)}$.

@@

%===============================================================================

TODO
Note: no face can have its edges glued together; prevented by fact
of alternating links

TODO just use Feasible Flow Theorem (see BS)

>>>>

%TODO: make a list of images of vectors, e.g.
%$\Theta_*(\ddd{\vphi_{\vec{e}}}) = -\ddd{\theta_e}$.

@@

To that end, let $L$ be the $(|E|-1)$-plane distribution on $\TTT$
tangent to the level sets of $\sum_{e\in E} \theta_e$.
The following proposition proves that we can deform
some circle patterns up to first order at a point:

\begin{proposition}
\label{p:point_lift}
Let $c \in \CCC$ be a
non-singular extended circle pattern on a graph
with no bigons or self-loops such that
there is no wide cycle of thin faces,
and each edge meets two distinct faces.
Let $K_c = \ker (\Phi_*|_c : T_c \CCC \to T_{\Phi(c)} \PPP)$.
Then $\Theta_*(K_c) = L_{\Theta(c)}$.

@@
